How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in \textit{Emys marmorata} (Testudinoidea, Emydidae).

### 2014-12-05 doc/draft_turtle.tex
methods, results mostly done. discussion points outlined. need to re-do unsupervised learning.

\subsection{Supervised learning}
% model selection/fit
Generally, as many PCs were included as predictors as possible for the best models of each of the classification schemes (Fig. \ref{fig:sel}). Note that the best random forest models were determined via recursive feature selection, so PCs were not included in order of percent variance explained. For both the LDA and multinomial logistic regression models, increasing model complexity increased cumulative percent variation necessary to best model the differing classification schemes (Fig. \ref{fig:sel}). The high variance in optimal number of PCs used as predictors between schemes may represent either ease of fitting that particular model or that the data was too noising to accurately fit any model beyond the simplest.

As part of fitting a random forest model, a ranking of variable importance is also determined. Interestingly, the the order of variable importance is not the same as the order of principal components (Fig. \ref{fig:var_imp}). This means that the principal components that best describe the differences between the various classes are not aligned with the principal components which describe the largest amount of variance. This result would be the case if variation between classes was extremely fine grained and not a part of the principal function or form of the plastron, which makes sense given that the plastron is involved in both protection and aquadynamics and not mate choice \citep{Germano2009,Holland1992,Lubcke2007,Rivera2008}.

@@

@@

\begin{table}[ht]
  \centering
  \include{comp_tab}
  \caption{CAPTION}
  \label{tab:comp}
\end{table}

% misfits
% need to make a table/graph to demonstrate this
Differences in mean shape between correctly and incorrectly classified observations from test set were frequently statistically significant, though there are exceptions. The frequency of these results, however, is important because this means that the different models are poor predictors of class membership. This may be because differences in plastron shape do not align with the any of the hypothesized classification schemes.

 % i have three tables for this. need to find a nice way of displaying these. also they require large amounts of edits
\begin{table}
  \begin{scriptsize}
    \include{rf_miss}

@@

Generalization and the unsupervised learning results are surprisingly congruous.

Plastron shape indicates no optimal classification scheme. A possible explanation for this that the genetic divergence associated with (sub)speciation is either not based on plastron morphology or local selective pressures due to hydrological regime overwhelming any possible morphological divergence.

>>>>

\subsection{Supervised learning}
% model selection/fit
AUC based model selection revealed some important patterns of variation and congruence between a given classification scheme and the actual data. Generally, as many PCs were included as predictors as possible for the best models of each of the classification schemes (Fig. \ref{fig:sel}). Note that the best random forest models were determined via recursive feature selection, so PCs were not included in order of percent variance explained. For both the LDA and multinomial logistic regression models, increasing model complexity increased cumulative percent variation necessary to best model the differing classification schemes (Fig. \ref{fig:sel}). That almost all models were as complex as possible indicates that the differences between the different groups within each classification scheme are very small.

As part of fitting a random forest model, a ranking of variable importance is also determined. Interestingly, the order of variable importance is not the same as the order of decreasing explained variance per principal component (Fig. \ref{fig:var_imp}). This means that the principal components that best describe the differences between the various classes are not aligned with the principal components which describe the largest amount of variance. This result would be the case if variation between classes was extremely fine grained and not a part of the principal function or form of the plastron, which makes sense given that the plastron is involved in both protection and aquadynamics and not mate choice \citep{Germano2009,Holland1992,Lubcke2007,Rivera2008}. This is congruous with the results of the AUC based model selection for the multinomial logistic regression and LDA models.

@@

% misfits
Differences in mean shape between correctly and incorrectly classified observations from test set were frequently statistically significant, though there are exceptions. The frequency of these results, however, is important because this means that the different models are poor predictors of class membership. This may be because differences in plastron shape do not align with the any of the hypothesized classification schemes.

The results of this study indicate that there is no clear grouping of plastra shapes in \textit{E. marmorata}.

The unsupervised learning results which indicate only a single group of observations being optimal is congruous with the results from the generalizations of the supervised learning models. The classification schemes used in the supervised learning models correspond, loosely, to unsupervised learning solutions with multiple groups. Because unsupervised learning solutions with multiple groups are poor descriptors of the observed variation, it is important to see this reinforced by the supervised learning results.

The results from fitting the various supervised learning models for each of the classification scheme generally shows that no one scheme is ``best.'' A possible explanation for this that the genetic divergence associated with (sub)speciation is either not based on plastron morphology or local selective pressures due to hydrological regime overwhelming any possible morphological divergence.

Both the low AUC values (\(< 0.9\)) and the significant difference between the correctly and incorrectly classified observations support the conclusion that none of the hypothesized classification schemes are good descriptors of the observed plastral variation.

@@

\begin{table}[ht]
  \centering
  \include{comp_tab}
  \caption{AUC values for the best model of each classification scheme for both the observed (training) data and the generalized (testing) data. Results from all three different supervised learning approaches are shown here. AUC values range between 0.5 and 1. }
  \label{tab:comp}
\end{table}

\begin{table}
  \begin{scriptsize}
    \include{rf_miss}

@@
